% This document provides the style to be used for a MSc Thesis at the
% Parallel and Distributed Systems group
\documentclass[11pt,twoside,a4paper,openright]{report}

% use babel for proper hyphenation
\usepackage[british]{babel}
% Graphics: like the DUT logo on the front cover
\usepackage[dvips]{graphicx}
%Enables [H] for figures.
\usepackage{float}
% FONT: times
\usepackage{times}
% for url's use "\url{http://www.google.com/}"
\usepackage{url}
% To allow margin adjustments
\usepackage{changepage}
% To allow listings.
\usepackage{listings}
% To inserts to do's
\usepackage{todonotes}
% Boxed verbatim for experiments
\usepackage{fancyvrb}
% Side by side packages
%\usepackage{subfigure}
% sub caption package
\usepackage{subcaption}
% bib in order of appearance
\bibliographystyle{ieeetr}

\newsavebox{\FVerbBox}
\newenvironment{FVerbatim}
 {\VerbatimEnvironment
  \begin{center}
  \begin{lrbox}{\FVerbBox}
  \begin{BVerbatim}}
 {\end{BVerbatim}
  \end{lrbox}
  \fbox{\usebox{\FVerbBox}}
  \end{center}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hoffset=1.63cm
\oddsidemargin=0in
\evensidemargin=0in
\textwidth=5in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% No identing
\setlength\parindent{0pt}

\pagestyle{empty}

% FRONTCOVER
\include{frontcover}

% EMPTY PAGE
% \cleardoublepage

\pagestyle{plain}

% TITLE PAGE: page i (hidden)
\include{titlepage}

% GRADUATION DATA AND ABSTRACT: pages ii and iii (hidden)
\include{graduationdata}

\pagenumbering{roman}
\setcounter{page}{4}

% EMPTY PAGE: page iv
% \cleardoublepage

% OPTIONAL QUOTATION: page v
%\include{quotation}
% EMPTY PAGE: page vi
%\cleardoublepage

% PREFACE: page v
\include{acknowledgements}
% EMPTY PAGE: page vi
% \clearpage

% TABLE OF CONTENTS: starting at page vii
\setcounter{tocdepth}{2}
\tableofcontents

% \cleardoublepage

\pagenumbering{arabic}
\setcounter{page}{1}

% CHAPTERS ... For instance: History/Prior Work, Design/Implementation, Experiments
% INTRODUCTION
\include{introduction/introduction}
%Problem description
\include{preliminaries/preliminaries}\section{BitTorrent protocol}
When sharing files by using the BitTorrent protocol, a peer that uploads parts of a file to another peer, is called a seeder.
A peer that is downloading a file of a seeder is called a leecher.
Any peer can be both a seeder and leecher at the same time, and join the network at any given time.

The ratio between the total data downloaded and uploaded is called the seeding ratio \cite{Cohen-bittorrent}.
The seeding ratio can be seen as an indication of the level of collaboration i.e. giving back resources to the network.

Seeding can be seen as an interaction between peers, where the seeder aids the leeching peer.
By utilizing the seeders upload bandwidth, the leeching peer can use his download bandwidth to download a file.
While there is a clear incentive for the leecher by downloading the desired file, there is none for the seeder.
Especially since the leecher has a little chance of becoming also be a seeder for the original seeder \cite{Lai-Incentives}.

Having peers actively and persistently contribute to the network will increase the network's health which in turn provides several benefits for all peers.
A more healthy network results in a higher availability of seeders and results in high download speeds.
It has been shown that private communities where the seeding ratio is high, provides better download conditions \cite{meulpolder-privatecommunities}.
In these private communities, trackers i.e. central components introduce peers to each other using the Tit-for-tat approach \cite{cohen-titfortat}.
The Tit-for-Tat approach is aiding peers who have aided you in the past,
The absence of trackers, which is often the case in public networks, results in free-riding \cite{Adar-Freeriding}.
A free-riding peer does not or gives little back to the network while receiving all benefits i.e. download without any restriction.
\todo{Explain the optimistic chocking approach.}BitTorrent applies a variation of the Tit-for-tat strategy, optimistic chocking, to combat this problem.
The Tit-for-Tat strategy is to only provide help to peers that return this help.
However, it has been shown that this approach is not effective in battling abuse \cite{Pouwelse-tribler}.
\include{problemDescription/problemDescription}
%Related work
\include{relatedWork/relatedWork}
%Experimentation
\include{experimentation/experimentation}
%Conclusion
\include{conclusion/conclusion}


% BIBLIOGRAPHY
%\bibliographystyle{bib/latex8}
\bibliography{bib/bibliography}

%\appendix

%\include{appendix_a}

\end{document}

